---
title: "Portfolio 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#loading packages
library(tidyverse)
library(jpeg)

```

1a)
```{r}
sleep = read.csv("sleepstudy.csv")


P372 = filter(sleep, Subject == 372)

Reaction372 = P372$Reaction

Days372 = P372$Days

#1a
Constant = rep(1,10)

#1b
#dot product
#Days372%*%Constant
sum(Days372*Constant)
```
#1C
Since the dot product is above 0 the two variables are positivly correlated.

### 1d

```{r}
#making the X matrix
X = matrix(c(Constant, Days372),ncol=2)

#making the linear regression manually
beta = solve( t(X)%*%X ) %*% t(X)%*%Reaction372

beta
```

### 1e

```{r}
#checking with the lm function.
summary(lm(Reaction372 ~0+X))
```
As can be seen they are identical.


### 1f

```{r}
#finding the mean of reaction time for participant 372
mean372 = mean(Days372)
newdays = Days372-mean372

X[,2] = newdays
X
#making the linear regression manually
beta1 = solve( t(X)%*%X ) %*% t(X)%*%Reaction372

#checking with the lm function.
summary(lm(Reaction372 ~0+X))

mean(Reaction372)
```
The intercept changed but the slope did not. The slope did not change because the difference between the predictor values did not change because they were subtracted from a constant value. The intercept changed to the mean of the reaction time.


```{r}
data = as.data.frame(newdays)

data$reaction = Reaction372

ggplot(data, aes(newdays, reaction))+geom_point()+geom_smooth(method = "lm", alpha = 0)+labs(x = "days", y = "Reaction time in mS")
```




## 2

### 2a
```{r}
matrix = readJPEG('portfolio_assignment2_matrices_data.jpg', native = FALSE)

ncol(matrix)
nrow(matrix)
min(matrix)
max(matrix)
mean(matrix)
```

### 2b
```{r}
image(matrix, col=grey(1:100/100))
#rotating it 90 degrees
rotate <- function(x) t(apply(x, 2, rev))
matrix<-rotate(matrix)

#convert to image

im = image(matrix, col=gray(1:100/100))
```


### 2c

```{r}
im = image(matrix2, col=gray(0))

matrix2 = matrix
```

### 2d
```{r}
#hat
matrix2[225:425,800:810] <- 1
#rectangle
matrix2[275:375,800:875] <- 1
# drawing the image
image(matrix2, col = gray(1:100/100))
```

### 2e
```{r}
whitehat = ifelse(matrix2 == 1,1,0)

image(matrix*whitehat, col=grey(1:100/100))

```

## 3

### 3a

```{r}
brain<-readJPEG('portfolio_assignment2_matrices_data2.jpg', native = FALSE)

brain = rotate(brain)

image(brain,col = grey(1:100/100))

mean(brain)

mask1 = ifelse(brain > mean(brain),1,0)

mask2 = ifelse(brain < 2.5*mean(brain),1,0)

mask3 = mask1*mask2

image(mask3*brain, col = grey(1:100/100))

sum(mask3)

```
there are 50004 pixels left with color


## 4a.

```{r}
q = matrix(c(3,1,1,2),ncol = 2)
v  = c(116,92)
solve(q)%*%v

28*3+32

28+2*32
2*116-92
-116+3*92
140/5
160/5
```

